Fast sparse period estimation

### 2015-04-14 poster/ITR_Poster.tex
minor edits to poster

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Fast sparse period estimation}
\author{Robby.~G.~McKilliam, I.~Vaughan.~L.~Clarkson \\ and Barry.~G.~Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported by Australian Research Council Discovery Project DP130102228}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{block}{Applications}
\begin{itemize}
\item Pulse interval estimation in electronic support%~\cite{Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}.
\item Synchronisation for telecommunications%~\cite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}.
\item Spike interval estimation in neurology%~\cite{Brillinger_spike_trains_1988}.
\item Detection of anomalies in internet traffic%~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
\end{itemize}
Connected with Diophantine approximation, lattice reduction, and the study of prime numbers.%~\cite{Clarkson_thesis,CaseySadler_primes_2013}.
\end{block}

@@

\begin{block}{The periodogram estimator}
An accurate estimator is $\hat{T} = 1/\hat{f}$ where $\hat{f}$ maximises the \emph{periodogram}
\[
I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2.
\]
over the interval $(1/\Tmax, 1/\Tmin)$.  
\begin{itemize}
\item Requires $O(N^2)$ operations to compute. 
\item Observations $y_1,\dots,y_N$ occur in the exponent preventing use of fast transforms.
\end{itemize}
\end{block}

@@

Compute quantized observations
\[
z_n =  \ell_n/q, \qquad n = 1,\dots,N
\] 
where $\ell_n = \round{qy_n} \in \ints$ and maximise
\[
I_z(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f z_n} }^2 = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{2\pi j f i/q} b_i }^2
\]
where $\ellmin$ and $\ellmax$ are the minimum and maximum value of $\ell_n$ and
\[
b_i= \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
\]
where $\#$ indicates the number of elements in a set.

@@

\begin{figure}[p] 
  \centering 
  \begin{tikzpicture} 
\begin{groupplot}[
    group style={ 
        group name=my plots,
        group size=1 by 2,
        %ylabels at=edge left,
        xlabels at=edge bottom,
        xticklabels at=edge bottom,
        vertical sep=0pt
    },
    ylabel style={at={(0.085,0.4)}},
    footnotesize,
    width=25cm,
    height=15.83cm,
    tickpos=left,
    %ytick align=outside,
    %xtick align=outside,
    xmode=log,
    ymode=log,
    xmin=11e-5,
    xmax=0.4
%    enlarge x limits=false 
]

@@

      \addplot[mark=none,color=black,dashed] table {../code/data/NormalisedLLSCLTN30geom10};

            \addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN30q1.5geom10};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN30q5.0geom10};
\addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN30geom10};
\addplot[mark=none,color=black] table {../code/data/PeriodogramCLTN30geom10}; 
 \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/NormalisedSamplingLLSN30geom10}; 
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN30geom10};
%      
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};

@@

\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Seperable line search,Least squares Theory,Periodogram Theory, Least squares,Periodogram}

@@

\begin{figure}
  \centering 
  \begin{tikzpicture}
%    \selectcolormodel{gray} 
    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.085,0.42)}},xlabel style={at={(0.5,0.13)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
      \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/LeastSquaresBenchmark};
      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
      \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newBenchmark};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
      %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
      \legend{Periodogram, Least squares, Seperable line search, Quantised periodogram $q=5$}
   \end{axis}
  \end{tikzpicture}  
  \caption{Computation time in milliseconds versus data length $N$.}\label{plot:benchmark}
\end{figure}

@@

\nocite{clarkson_estimate_period_pulse_train_1996,Sidiropoulos2005,5585849,Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}

>>>>

%%% title, author & contact %%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Fast sparse period estimation}
\author{Robby~G.~McKilliam, I.~Vaughan~L.~Clarkson \\ and Barry~G.~Quinn}
\newcommand{\email}{robby.mckilliam@unisa.edu.au}
\newcommand{\worksupportedby}{Supported by Australian Research Council Discovery Project DP130102228}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{block}{Applications}
\begin{itemize}
\item Pulse interval estimation in electronic support.%~\cite{Clarkson_thesis,clarkson_estimate_period_pulse_train_1996,Hauochan_pri_2012}.
\item Synchronisation for telecommunications.%~\cite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,5621928}.
\item Spike interval estimation in neurology.%~\cite{Brillinger_spike_trains_1988}.
\item Detection of anomalies in internet traffic.%~\cite{He_detecting_periodic_patterns_in_internet_2006,5585849,5947313}.
\end{itemize}
Connected with Diophantine approximation, lattice reduction, and the study of prime numbers.%~\cite{Clarkson_thesis,CaseySadler_primes_2013}.
\end{block}

@@

\begin{block}{The periodogram estimator}
An accurate estimator is $\widehat{T} = 1/\widehat{f}$ where $\widehat{f}$ maximises the \emph{periodogram}
\[
I_y(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f y_n} }^2.
\]
over the interval $(1/\Tmax, 1/\Tmin)$.  
\begin{itemize}
\item Requires $O(N^2)$ operations to compute. 
\item Observations $y_1,\dots,y_N$ occur in the exponent preventing use of fast transforms.
\end{itemize}
\end{block}

@@

Compute quantised observations
\[
z_n =  \ell_n/q, \qquad n = 1,\dots,N
\] 
where $\ell_n = \round{qy_n} \in \ints$ and maximise
\[
I_z(f) = \babs{ \sum_{n=1}^N e^{ 2\pi j f z_n} }^2 = \babs{ \sum_{i = \ellmin}^{\ellmax} e^{2\pi j f i/q} b_i }^2
\]
where $\ellmin$ and $\ellmax$ are the minimum and maximum of $\ell_1,\dots,\ell_n$ and
\[
b_i= \#\{ n \mid \ell_n = i, n = 1,\dots, N\}
\]
where $\#$ indicates the number of elements in a set.

@@

\begin{figure}[p] 
  \centering 
  \begin{tikzpicture} 
\begin{groupplot}[
    group style={ 
        group name=my plots,
        group size=1 by 2,
        %ylabels at=edge left,
        xlabels at=edge bottom,
        xticklabels at=edge bottom,
        vertical sep=0pt
    },
    ylabel style={at={(0.085,0.4)}},
    footnotesize,
    width=25cm,
    height=16cm,
    tickpos=left,
    %ytick align=outside,
    %xtick align=outside,
    xmode=log,
    ymode=log,
    xmin=11e-5,
    xmax=0.4
%    enlarge x limits=false 
]

@@

      \addplot[mark=none,color=black,dashed,thick] table {../code/data/NormalisedLLSCLTN30geom10};

            \addplot[mark=*,mark options={scale=1.5},color=cyan] table {../code/data/QuantisedPeriodogramFFTN30q1.5geom10};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramFFTN30q5.0geom10};
\addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newN30geom10};
\addplot[mark=none,color=black,thick] table {../code/data/PeriodogramCLTN30geom10}; 
 \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/NormalisedSamplingLLSN30geom10}; 
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramN30geom10};
%      
  %     \addplot[mark=square,only marks,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpN30geom10};

@@

\legend{Quantised periodogram $q=1.5$, Quantised periodogram $q=5$,Separable line search,Least squares Theory,Periodogram Theory, Least squares,Periodogram}

@@

\begin{figure}
  \centering 
  \begin{tikzpicture}
%    \selectcolormodel{gray} 
    \begin{axis}[font=\footnotesize,xmode=log,ymode=log,height=15.25cm,width=25cm,xlabel={$N$},ylabel={time (ms)},ylabel style={at={(0.085,0.42)}},xlabel style={at={(0.5,0.13)}}, legend style={draw=none,fill=none,legend pos=north west,cells={anchor=west},font=\footnotesize},xmin=7,xmax=36000]
      \addplot[mark=*,mark options={scale=1.75},color=red] table {../code/data/PeriodogramBenchmark};
      \addplot[mark=o,mark options={scale=2.5},color=black, thick] table {../code/data/LeastSquaresBenchmark};
      %\addplot[mark=triangle*,mark options={scale=3},color=cyan] table {../code/data/QuantisedPeriodogramChirpZq4.0Benchmark};
      %\addplot[mark=square,color=black,mark options={solid,fill=black,scale=0.8}] table {code/data/SLS2novlpBenchmark};
      \addplot[mark=triangle*,mark options={scale=3},color=blue] table {../code/data/SLS2newBenchmark};
      \addplot[mark=square*,mark options={scale=2.5},color=green] table {../code/data/QuantisedPeriodogramq4.0Benchmark};
      %\legend{Periodogram, Least squares, Fast Fourier transform, Chirp z-transform, SLS-new}
      \legend{Periodogram, Least squares, Separable line search, Quantised periodogram $q=5$}
   \end{axis}
  \end{tikzpicture}  
  \caption{Computation time in milliseconds versus data length $N$.}\label{plot:benchmark}
\end{figure}

@@

\nocite{Fogel1989_bit_synch_zero_crossings,Sidiropoulos2005,Clarkson2007,Quinn_sparse_noisy_SSP_2012,Quinn20013asilomar_period_est}
